---
title: "Carnivores"
output: html_document
date: '2022-07-20'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Loading the data
```{r}
speciesE <- read.delim("speciesE.txt", header=T)
rownames(speciesE) <- speciesE$species_name


#defining var class
speciesE$diet.1 <- as.factor(speciesE$diet.1)
speciesE$extant <- as.factor(speciesE$extant)
speciesE$loctype <- as.factor(speciesE$loctype)
speciesE$diet.cat <- as.factor(speciesE$diet.cat)


#remove NA level from diet
speciesE$diet.1 <- factor(speciesE$diet.1, exclude="")
speciesE$diet.cat <- factor(speciesE$diet.cat, exclude="")
speciesE$diet.new <- factor(speciesE$diet.new, exclude="")
```

##Marking extinct vs extant
```{r}
extinct_large_mammals <- read.delim("extinct_large_mammals.txt", header=F)
extant_large_mammals <- read.delim("extant_large_mammals.txt", header=F)

#check for missing species
extant <- match(speciesE$species, extant_large_mammals$V1)
extinct <- match(speciesE$species, extinct_large_mammals$V1)

sum(names(table(speciesE$species)) %in% names(table(extinct_large_mammals)))  #count common species between ecoreg and Extinct list = 19
sum(names(table(speciesE$species)) %in% names(table(extant_large_mammals)))   #count common species between ecoreg and Extant list = 95

which(is.na(extant)&(is.na(extinct))) #check for double NAs
check <- cbind (extant, extinct) #check for double NAs

#match extant species in the dataset and mark them with 0 = extinct and 1 = extant

m <- match(speciesE$species, extant_large_mammals$V1)
speciesE$extant <- m
speciesE$extant[!is.na(speciesE$extant)] <- 1
speciesE$extant[is.na(speciesE$extant)] <- 0
```


##Add data from other datasets
```{r}

#HELBSTAB
#-------------------------------------
  

#Split the genus from family
speciesH <- array()
s <- strsplit(X_helbstab_journal_pone_0261185_s005$`Species name`,'_')

  for (i in 1:nrow (X_helbstab_journal_pone_0261185_s005)) {
  speciesH[i] = paste(s[[i]][1], s[[i]][2])
 }

#X_helbstab_journal_pone_0261185_s005$species1 <- strsplit(X_helbstab_journal_pone_0261185_s005$`Species name`,'_')[[1]][1]

#Count the common species with Heldstab
sum(names(table(speciesH)) %in% names(table(species$V1)))


#get the brains
m1 <- match(speciesE$species, speciesH)
speciesE$brain <- X_helbstab_journal_pone_0261185_s005$`Log ECV`[m1]
speciesE$EQ <- X_helbstab_journal_pone_0261185_s005$`Encephalization quotient (EQ)`[m1]

#write to main data set
write.table(speciesE, "speciesE.txt", sep = "\t", row.names = FALSE)


#NOONAN PC and Litter Size
#-----------------------------------

noonan_tables1 <- as.data.frame(noonan_tables1)
data <- noonan_tables1[, 5:15] #activity vars
rownames(data) <- paste(noonan_tables1[,'Genus'], noonan_tables1[,'Species'])
pc <- princomp(na.omit(data))
pc_noonan <- pc$scores[,1]
names(pc_noonan) <- rownames(pc$scores)
w <- match(speciesE$species, names(pc_noonan))
speciesE$pc <- pc_noonan[w]

#Also add litter size
#speciesE <- speciesE2[1:114,]


#BEN BMR
#--------------------------------------

#Adding BMR (from two diff columns same df)

w <- match(speciesE$species, All_Carnivore_Species_BMRs$species_name) #match names
w1 <- match(speciesE$species, All_Carnivore_Species_BMRs$synonym) #match synonyms
speciesE$BMR <- All_Carnivore_Species_BMRs$BMR_ml_O2_h[w]
speciesE$BMR1 <- All_Carnivore_Species_BMRs$BMR_ml_O2_h[w1]
speciesE$BMR[is.na(speciesE$BMR)] = data$BMR1[!is.na(speciesE$BMR1)]

```

##Diet re-cat
```{r}

#pure carn
#invert carn
#frug carn
#frug invert
#browser primary

speciesE1 <- speciesE
speciesE1$diet.new <- ifelse((speciesE$diet.1 == "carnivore") & (speciesE$diet.2 == ""), "carnivore","")
speciesE1$diet.new <- ifelse((speciesE$diet.1 == "invertivore") & (speciesE$diet.2 == ""), "invertivore",speciesE1$diet.new)
#speciesE1$diet.new <- ifelse((speciesE$diet.1 == "browser") & (speciesE$diet.2 == ""), "browser",speciesE1$diet.new)
speciesE1$diet.new <- ifelse((speciesE$diet.1 == "browser"), "browser",speciesE1$diet.new)
speciesE1$diet.new <- ifelse((speciesE$diet.1 == "frugivore") & (speciesE$diet.2 == ""), "frugivore",speciesE1$diet.new)
speciesE1$diet.new <- ifelse((speciesE$diet.1 == "piscivore"), "piscivore",speciesE1$diet.new)

speciesE1$diet.new <- ifelse((speciesE1$diet.1 == "invertivore") & (speciesE1$diet.2 == "carnivore"), "carninvert", speciesE1$diet.new)
speciesE1$diet.new <- ifelse((speciesE1$diet.1 == "carnivore") & (speciesE1$diet.2 == "invertivore"), "carninvert", speciesE1$diet.new)


speciesE1$diet.new <- ifelse((speciesE1$diet.1 == "carnivore") & (speciesE1$diet.2 == "frugivore"), "carnifrug", speciesE1$diet.new)
speciesE1$diet.new <- ifelse((speciesE1$diet.1 == "carnivore") & (speciesE1$diet.2 == "granivore"), "carnifrug", speciesE1$diet.new)
speciesE1$diet.new <- ifelse((speciesE1$diet.1 == "invertivore") & (speciesE1$diet.2 == "frugivore"), "invertivore", speciesE1$diet.new)



speciesE1$diet.new <- ifelse((speciesE1$diet.1 == "frugivore") & (speciesE1$diet.2 == "carnivore"), "frugcarn", speciesE1$diet.new)
speciesE1$diet.new <- ifelse((speciesE1$diet.1 == "frugivore") & (speciesE1$diet.2 == "invertivore"), "frugivore", speciesE1$diet.new)

speciesE$diet.new <- speciesE1$diet.new


```


#Wrangle, explore and investigate
```{r}
require(easystats)
require(psych)
describe_distribution(speciesE)
multi.hist(speciesE[,sapply(speciesE, is.numeric)], global = F) 

speciesE$Log.mass.g <- log(speciesE$mass.g)
speciesE$Log.home.range <- log(speciesE$home.range)
speciesE$Log.population.density <- log(speciesE$population.density)
speciesE$Log.BMR <- log(speciesE$BMR)


data_filter(data1, extant == 0 & is.na(diet.new))


require(lattice)
dotplot(factor(speciesE$extant) ~ log(speciesE$mass.g))
```


#Misingness (and imputation)
```{r}
require(naniar)
require(mice)

gg_miss_var(speciesE)

```

#Tree
```{r}
require(phytools)
require(BAMMtools)


treeF <- read.tree("Complete_phylogeny.nex")
rownames(speciesE) <- speciesE$species_

#Adding the 3 missing species (length = 0.5 on the branch - except for Conepatus = 0 from the mrca)

for (i in 1:length(tree))	{
tip <- "Martes_caurina"
sister <- "Martes_americana"

tree[i] <- c(bind.tip(tree[[i]],tip,where=which(tree[[i]]$tip.label==sister),
    position=0.5*tree[[i]]$edge.length[which(tree[[i]]$edge[,2]==
    which(tree[[i]]$tip.label==sister))])) 


tip <- "Conepatus_robustus"
sister <- "Conepatus_chinga"
sister2 <- "Conepatus_leuconotus" 

tree[i] <- c(bind.tip(tree[[i]],tip,where=getmrca(tree[[i]], sister, sister2),
    position=0*tree[[i]]$edge.length[getmrca(tree[[i]], sister, sister2)])) 

tip <- "Felis_lybica"
sister <- "Felis_silvestris"

tree[i] <- c(bind.tip(tree[[i]],tip,where=which(tree[[i]]$tip.label==sister),
    position=0.5*tree[[i]]$edge.length[which(tree[[i]]$edge[,2]==
    which(tree[[i]]$tip.label==sister))])) }


grep("Martes", tree[[7]]$tip.label, value = TRUE)


clean <- clean.data(speciesE, tree)
tree1 <- clean$tree[[1]] #we can use all of the trees in mcmcglmm
data1 <- clean$data

```



```{r}
#adding missing species loop (bad solution with tree extraction!)

#extract all trees
treeNew <- list()
class(treeNew) <- "multiPhylo"


for(ntrees in 1:length(tree)) {
    assign(paste0("treeX", ntrees), tree[[ntrees]])
}

for (i in 1:length(tree))	{
tip <- "Martes_caurina"
sister <- "Martes_americana"

treeNew[i] <- c(bind.tip(get(paste0("treeX",i)),tip,where=which((get(paste0("treeX",i)))$tip.label==sister),
    position=0.5*(get(paste0("treeX",i)))$edge.length[which((get(paste0("treeX",i)))$edge[,2]==
    which((get(paste0("treeX",i)))$tip.label==sister))])) }
    
for(ntrees in 1:length(tree)) {
    assign(paste0("treeX", ntrees), treeNew[[ntrees]])
}


for (i in 1:length(tree))	{
tip <- "Felis_lybica"
sister <- "Felis_silvestris"
treeNew[i] <- c(bind.tip(get(paste0("treeX",i)),tip,where=which((get(paste0("treeX",i)))$tip.label==sister),
    position=0.5*(get(paste0("treeX",i)))$edge.length[which((get(paste0("treeX",i)))$edge[,2]==
    which((get(paste0("treeX",i)))$tip.label==sister))])) }

for(ntrees in 1:length(tree)) {
    assign(paste0("treeX", ntrees), treeNew[[ntrees]])
}

for (i in 1:length(tree))	{
tip <- "Conepatus_robustus"
sister <- "Conepatus_leuconotus"
treeNew[i] <- c(bind.tip(get(paste0("treeX",i)),tip,where=which((get(paste0("treeX",i)))$tip.label==sister),
    position=0.5*(get(paste0("treeX",i)))$edge.length[which((get(paste0("treeX",i)))$edge[,2]==
    which((get(paste0("treeX",i)))$tip.label==sister))])) }


rm(list=ls(pattern="treeX"))
```


#Load final tree and data
```{r}
require(phytools)
tree <- read.tree("CompletePhyCarn.txt")
data <- speciesE
#data <- speciesE
rownames(data) <- data$species_name




require(wesanderson)
pal <- wes_palette("Darjeeling1", 13, type = "continuous")



```

#Analyses - gls;phyloglm;rstanarm
```{r}

require(phylolm)
require(mulTree)
require(rstanarm)


#Non-phylo gls
a <- glm(speciesE$extant ~ speciesE$diet.1, family = binomial)
summary(a)

#RELEVEL for ancova relevel(diet.1, ref = "carnivore")

#timetree phy
tree2 <- read.newick("spcs.nwk")
clean.data(speciesE, tree)


aa <- phyloglm(extant ~ log(mass.g) + loctype, speciesE, tree) #scansorial+
aa <- phyloglm(extant ~ median.abundance + loctype, speciesE, tree) #terrestrial-
aa <- phyloglm(extant ~ diet.1, speciesE, tree) #carnivore+
aa <- phyloglm(extant ~ median.abundance + log(mass.g) + loctype, speciesE, tree) #median abd+, mass-

#phylacine phy
treeF <- read.nexus("Complete_phylogeny.nex")
rownames(speciesE) <- speciesE$species_
clean <- clean.data(speciesE, treeF)
tree1 <- clean$tree[1]$UNTITLED #we can use all of the trees in mcmcglmm
data1 <- clean$data

#check the diet var!
#data1$diet.new <- as.factor(data1$diet.new)
#data1$diet.new <- factor(data1$diet.new, exclude="")

phyloglm(extant ~ log(BMR) + diet.1, data1, tree1, method = c("logistic_MPLE"), btol = 100, boot=10) # also only diet
phyloglm(extant ~ median.abundance + log(mass.g) + loctype, data1, tree1, method = c("logistic_MPLE"), btol = 30, boot = 100) # mass?

#rstanarm

fit_rstanarm <- stan_glm(
    extant ~ loctype,
    data = speciesE,
    family = "binomial"
)

summary (fit_rstanarm)
```


```{r}

print.pval <- function(x, digits = 4) {
    sig_level <- cut(x,
                     breaks = c(-Inf, 0, 0.0001, 0.001, 0.01, 0.05, Inf),
                     labels = c("", "****", "***", "**", "*", ""))
    print(paste0(signif(x, digits = digits), sig_level), quote = FALSE)
    invisible(x)
}


regress <- function (multree, data, model, from, to) {
    a = 0
for (i in from:to){
    temptree <- multree[[i]]
    s1 <- phyloglm(model, data, temptree, method = c("logistic_MPLE"), btol=30)
    s2 <- as.data.frame(summary(s1)$coefficients[,4])
    asd <- as.numeric(s2[[1]])
  #    for (t in 1:length(s2[[1]]))  {
  #     asd <- if(s2[[1]][t]<0.06){sprintf(">>>%f<<<", s2[[1]][t])}
  #    }
      if(any(asd<0.06)){print.pval(asd)} 
      if(any(s2<0.06)){a = a+1
      cat(asd,"\n", " N Sig models:", a, " out of", i-1, "", round((a/(to-from))*100, 2), "% " ,"\n")
  
      }
     
                  
        #print(summary(s1)$coefficients[,4])
        #print (summary(s1))
                  }
    cat(paste0("\n Predictors:"))
    cat(paste0("\n", "", row.names(s2)[-1]), "\n")
    cat(paste0("\n ==================== \n", " Total models ", to - from, "\n Significant ", a, "\n ", round((a/(to-from))*100, 2), "% ", "\n ==================== \n"))
  
                                                }

regress(tree, data, extant ~ diet.new, 1, 12)


                                    
```

#Plots
```{r}
data <- as.data.frame(speciesE$extant)
rownames(data) <- speciesE$species_
data$loc <- speciesE$loctype
data$loc <- as.factor(data$loc)
data$data <- as.factor(data$data)

require(ggtree)
require(ggplot)


circ <- ggtree(tree, layout = "circular")
p1 <- gheatmap(circ, data, offset=.8, width=.2,
               colnames_angle=95, colnames_offset_y = .25) +
    scale_fill_viridis_d(option="D", name="discrete\nvalue") + geom_tippoint(aes(color = c("a", "b")))

p2 <- ggplot(data, aes(x=diet, y=tree1$tip.label)) + 
    geom_tile(aes(fill=`clean$data$extant`)) + scale_fill_viridis_d() + 
    theme_minimal() + xlab(NULL) + ylab(NULL)

#+ geom_tippoint(aes(color = c("a", "b")))


circ <- ggtree(tree1, layout = "circular") + geom_tiplab(offset=15)
g <- ggtree(tree1) + geom_tiplab(align=TRUE) + hexpand(.5)
p1 <- gheatmap(circ, data[,c("diet.1", "diet.cat", "diet.new")], offset=.8, width=.2,colnames_angle=95, colnames_offset_y = .25)


grep("robustus", b, value = TRUE)


#heatmap using ggtree
require(ggtree)
p1 <- ggtree(tree[[1]]) + 
    xlim(0, 150) +
    geom_tiplab(size=4, offset = 35)

p2 <-  gheatmap (p1, data[,c("diet.1", "diet.cat", "diet.new")], offset=-0.5, width=0.5, low=(values=pal[1]), high=(values=pal[2]), colnames_position = "top", font.size=4) + scale_fill_manual(values = pal)

#p2 <-  gheatmap (p1, data[,c(26, 20, 14)], offset=-0.5, width=0.5, low="bisque", high="darkred", colnames_position = "top", font.size=4) + scale_fill_manual(values=c("red", "black", "blue", "green", "yellow", "purple", "pink", "cyan", "brown", "coral"))

#phylo.heatmap(tree,X,standardize=TRUE)

```


#SCAT
```{r}
s2 = read.delim('all_scats.txt')
s2[is.na(s2)] = 0
s2 = s2[,33:54]
s2 = t(t(s2) / colSums(s2))

s = read.delim('all_scats.txt')
sp = paste(s[,29],s[,30])
s = s[,33:54]
s[is.na(s)] = 0
s = s / rowSums(s)

d = matrix(nrow=nrow(s),ncol=6,dimnames=list(sp,c('vert','invert','frug','plant','aquatic','fungus')))

d[,1] = rowSums(s[,1:4])
d[,2] = rowSums(s[,7:12])
d[,3] = rowSums(s[,14:17])
d[,4] = rowSums(s[,18:20])
d[,5] = rowSums(s[,c(5,6,13)])
d[,6] = rowSums(s[,21:22])

plot(hclust(dist(d)),labels=sp,cex=0.5)

k = kmeans(d,centers=5)$clust

sort(k)

v = array()

for (i in 1:20)	{
	v[i] = (kmeans(d,centers=i)$tot.withinss / kmeans(d,centers=i)$totss)
}

plot(v,type='l')


barplot(t(d[order(k),]),col=hsv(h=0:5 / 6))


#add to main datatset
dc.match <- match(speciesE$species, names(k))
speciesE$diet.cat <- k1$k[dc.match]


#ed = vert, yellow = invert, green = frug, cyan = plant, blue = aquatic, ? = fungus


#Indeces
require(vegan)

s2 = read.delim('all_scats.txt')
scat = s2[,33:54]
scat[is.na(scat)] = 0
scat$species <- paste(s2[,'genus'], s2[,'species'])


#semi-manual with vegan
for (i in 1:length(scat$species))	{
    assign(paste0("n",i), as.numeric(scat[i,1:22]))
}

for (i in 1:length(scat$species))	{
     scat$shannon[i] <- diversity(get(paste0("n",i)), index = "shannon")
}

for (i in 1:length(scat$species))	{
     scat$simpson[i] <- diversity(get(paste0("n",i)), index = "simpson")
}

rm(list=ls(pattern="n"))


#OR EASIER

s2 = read.delim('all_scats.txt')
scat = s2[,33:54]
scat[is.na(scat)] = 0
scat$species <- paste(s2[,'genus'], s2[,'species'])
scat.species <- aggregate(. ~ species, data=scat, FUN=mean)
rownames(scat.species) <- scat.species$species
scat.species$species <- NULL
scat.species$shannon <- diversity(scat.species, index="shannon")
scat.species$simpson <- diversity(scat.species, index="simpson")

scat.match <- match(speciesE$species, row.names(scat.species))
speciesE$shannon <- scat.species$shannon[scat.match]
speciesE$simpson <- scat.species$simpson[scat.match]

#s2 = read.delim('all_scats.txt')
#scat = s2[,33:54]
#scat[is.na(scat)] = 0
#scat$species <- paste(s2[,'genus'], s2[,'species'])

#Using SYNCSA for Rao's QE
require(SYNCSA)
scat.species$shannon <- NULL
scat.species$simpson <- NULL
rownames(scat.species) <- scat.species$species
scat.species$species <- NULL
r1 <- rao.diversity(scat.species)
scat.match <- match(speciesE$species, rownames(scat.species))
speciesE$raoS <- r1$Simpson[scat.match]


write.table(speciesE, "speciesE.txt", sep = "\t", row.names = FALSE)



#scat$sums <- rowSums(scat[1:22])


#We only have diet on 2 extinct species: Martes caurina AND Felis lybica


#-----------------------------
#manual calc

#n = pop
#N = sum(row)

for (i in 1:length(s2$species))	{
    assign(paste0("n",i), as.numeric(scat[i,1:22]))
}

N = rowSums(scat[1:22])


#p = n/N
for (i in 1:length(s2$species))	{
    assign(paste0("p",i),  get(paste0("n",i))/N[i])
}
  

for (i in 1:length(s2$species))	{
    scat$shannon[i,] <- diversity(get(paste0("n",i)))
}

H = -sum(p*log(p))  

rm(list=ls(pattern="n"))

#------------------------------------

```



###############3


species -> register
register <- sample
registerX$species_full <- trimws(registerX$species_full, which = c("right"))


rnum <- array(dim=245, dimnames = list(as.character(nos)), data = 1:245) 



for (i in nos){
    w = which(registerX[,'sample.no'] == i)
    samplesX[rnum[as.character(i)], "countT"] = sum(registerX[w, 'count'])
}






change 0s to 0.00001s
primatedata[which(primatedata[,14]==0, arr.ind=TRUE), 14] <- 0.00001


put the new values in a dataframe

df <- data.frame(x,y,z)







