---
title: "Carnivores"
output: html_document
date: '2022-07-20'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#1. Load data
```{r}
require(psych)

speciesE <- read.delim("speciesE.txt", header=T)
rownames(speciesE) <- speciesE$species_name


#defining var class
speciesE$extant <- as.factor(speciesE$extant)
speciesE$loctype <- as.factor(speciesE$loctype)
speciesE$diet.cat <- as.factor(speciesE$diet.cat)


#remove NA level from diet
#speciesE$diet.cat <- factor(speciesE$diet.cat, exclude="")
#speciesE$diet.new <- factor(speciesE$diet.new, exclude="")

#Log vars
require(VGAM)

speciesE$simpson[23] <- NA
speciesE$simpson[78] <- NA
speciesE$simpson[80] <- NA

speciesE$mass.g <- log(speciesE$mass.g)
speciesE$home.range <- log(speciesE$home.range)
speciesE$BMR <- log(speciesE$BMR)
speciesE$simpson <- log(1/speciesE$simpson)


yj <- function (y) {((y+1)^0.0001 - 1) / 0.0001}
a <- (which(!is.na(speciesE$abundance)))
a1 <- yj(which(!is.na(speciesE$abundance)))
speciesE$abundance[a] <- a1[a] #transform Yeo–Johnson


#do we need these?
speciesE$mean.abundance <- NULL #should use median? - recode!
speciesE$diet.new <- NULL

multi.hist(speciesE[,sapply(speciesE, is.numeric)], global = F) 
```

##Marking extinct vs extant
```{r}
extinct_large_mammals <- read.delim("extinct_large_mammals.txt", header=F)
extant_large_mammals <- read.delim("extant_large_mammals.txt", header=F)

#check for missing species
extant <- match(speciesE$species, extant_large_mammals$V1)
extinct <- match(speciesE$species, extinct_large_mammals$V1)

sum(names(table(speciesE$species)) %in% names(table(extinct_large_mammals)))  #count common species between ecoreg and Extinct list = 19
sum(names(table(speciesE$species)) %in% names(table(extant_large_mammals)))   #count common species between ecoreg and Extant list = 95

which(is.na(extant)&(is.na(extinct))) #check for double NAs
check <- cbind (extant, extinct) #check for double NAs

#match extant species in the dataset and mark them with 0 = extinct and 1 = extant

m <- match(speciesE$species, extant_large_mammals$V1)
speciesE$extant <- m
speciesE$extant[!is.na(speciesE$extant)] <- 1
speciesE$extant[is.na(speciesE$extant)] <- 0
```


#Add data from other datasets
```{r}

#HELBSTAB
#-------------------------------------
  

#Split the genus from family
speciesH <- array()
s <- strsplit(X_helbstab_journal_pone_0261185_s005$`Species name`,'_')

  for (i in 1:nrow (X_helbstab_journal_pone_0261185_s005)) {
  speciesH[i] = paste(s[[i]][1], s[[i]][2])
 }

#X_helbstab_journal_pone_0261185_s005$species1 <- strsplit(X_helbstab_journal_pone_0261185_s005$`Species name`,'_')[[1]][1]

#Count the common species with Heldstab
sum(names(table(speciesH)) %in% names(table(species$V1)))


#get the brains
m1 <- match(speciesE$species, speciesH)
speciesE$brain <- X_helbstab_journal_pone_0261185_s005$`Log ECV`[m1]
speciesE$EQ <- X_helbstab_journal_pone_0261185_s005$`Encephalization quotient (EQ)`[m1]

#write to main data set
write.table(speciesE, "speciesE.txt", sep = "\t", row.names = FALSE)


#NOONAN PC and Litter Size
#-----------------------------------

noonan_tables1 <- as.data.frame(noonan_tables1)
data <- noonan_tables1[, 5:15] #activity vars
rownames(data) <- paste(noonan_tables1[,'Genus'], noonan_tables1[,'Species'])
pc <- princomp(na.omit(data))
pc_noonan <- pc$scores[,1]
names(pc_noonan) <- rownames(pc$scores)
w <- match(speciesE$species, names(pc_noonan))
speciesE$pc <- pc_noonan[w]

#Also add litter size
#speciesE <- speciesE2[1:114,]


#BEN BMR
#--------------------------------------

#Adding BMR (from two diff columns same df)

w <- match(speciesE$species, All_Carnivore_Species_BMRs$species_name) #match names
w1 <- match(speciesE$species, All_Carnivore_Species_BMRs$synonym) #match synonyms
speciesE$BMR <- All_Carnivore_Species_BMRs$BMR_ml_O2_h[w]
speciesE$BMR1 <- All_Carnivore_Species_BMRs$BMR_ml_O2_h[w1]
speciesE$BMR[is.na(speciesE$BMR)] = data$BMR1[!is.na(speciesE$BMR1)]

```

##Diet re-cat
```{r}

#pure carn
#invert carn
#frug carn
#frug invert
#browser primary

speciesE1 <- speciesE
speciesE1$diet.new <- ifelse((speciesE$diet.1 == "carnivore") & (speciesE$diet.2 == ""), "carnivore","")
speciesE1$diet.new <- ifelse((speciesE$diet.1 == "invertivore") & (speciesE$diet.2 == ""), "invertivore",speciesE1$diet.new)
#speciesE1$diet.new <- ifelse((speciesE$diet.1 == "browser") & (speciesE$diet.2 == ""), "browser",speciesE1$diet.new)
speciesE1$diet.new <- ifelse((speciesE$diet.1 == "browser"), "browser",speciesE1$diet.new)
speciesE1$diet.new <- ifelse((speciesE$diet.1 == "frugivore") & (speciesE$diet.2 == ""), "frugivore",speciesE1$diet.new)
speciesE1$diet.new <- ifelse((speciesE$diet.1 == "piscivore"), "piscivore",speciesE1$diet.new)

speciesE1$diet.new <- ifelse((speciesE1$diet.1 == "invertivore") & (speciesE1$diet.2 == "carnivore"), "carninvert", speciesE1$diet.new)
speciesE1$diet.new <- ifelse((speciesE1$diet.1 == "carnivore") & (speciesE1$diet.2 == "invertivore"), "carninvert", speciesE1$diet.new)


speciesE1$diet.new <- ifelse((speciesE1$diet.1 == "carnivore") & (speciesE1$diet.2 == "frugivore"), "carnifrug", speciesE1$diet.new)
speciesE1$diet.new <- ifelse((speciesE1$diet.1 == "carnivore") & (speciesE1$diet.2 == "granivore"), "carnifrug", speciesE1$diet.new)
speciesE1$diet.new <- ifelse((speciesE1$diet.1 == "invertivore") & (speciesE1$diet.2 == "frugivore"), "invertivore", speciesE1$diet.new)



speciesE1$diet.new <- ifelse((speciesE1$diet.1 == "frugivore") & (speciesE1$diet.2 == "carnivore"), "frugcarn", speciesE1$diet.new)
speciesE1$diet.new <- ifelse((speciesE1$diet.1 == "frugivore") & (speciesE1$diet.2 == "invertivore"), "frugivore", speciesE1$diet.new)

speciesE$diet.new <- speciesE1$diet.new


```

##SCAT
```{r}
s2 = read.delim('all_scats.txt')
s2[is.na(s2)] = 0
s2 = s2[,33:54]
s2 = t(t(s2) / colSums(s2))

s = read.delim('all_scats.txt')
sp = paste(s[,29],s[,30])
s = s[,33:54]
s[is.na(s)] = 0
s = s / rowSums(s)

d = matrix(nrow=nrow(s),ncol=6,dimnames=list(sp,c('vert','invert','frug','plant','aquatic','fungus')))

d[,1] = rowSums(s[,1:4])
d[,2] = rowSums(s[,7:12])
d[,3] = rowSums(s[,14:17])
d[,4] = rowSums(s[,18:20])
d[,5] = rowSums(s[,c(5,6,13)])
d[,6] = rowSums(s[,21:22])

plot(hclust(dist(d)),labels=sp,cex=0.5)

k = kmeans(d,centers=5)$clust

sort(k)

v = array()

for (i in 1:20)	{
	v[i] = (kmeans(d,centers=i)$tot.withinss / kmeans(d,centers=i)$totss)
}

plot(v,type='l')


barplot(t(d[order(k),]),col=hsv(h=0:5 / 6))


#add to main datatset
dc.match <- match(speciesE$species, names(k))
speciesE$diet.cat <- k1$k[dc.match]


#ed = vert, yellow = invert, green = frug, cyan = plant, blue = aquatic, ? = fungus


#Indeces
require(vegan)

s2 = read.delim('all_scats.txt')
scat = s2[,33:54]
scat[is.na(scat)] = 0
scat$species <- paste(s2[,'genus'], s2[,'species'])


#semi-manual with vegan (skip)
for (i in 1:length(scat$species))	{
    assign(paste0("n",i), as.numeric(scat[i,1:22]))
}

for (i in 1:length(scat$species))	{
     scat$shannon[i] <- diversity(get(paste0("n",i)), index = "shannon")
}

for (i in 1:length(scat$species))	{
     scat$simpson[i] <- diversity(get(paste0("n",i)), index = "simpson")
}

rm(list=ls(pattern="n"))


#OR EASIER

s2 = read.delim('all_scats.txt')
scat = s2[,33:54]
scat[is.na(scat)] = 0
scat$species <- paste(s2[,'genus'], s2[,'species'])
scat.species <- aggregate(. ~ species, data=scat, FUN=mean)
rownames(scat.species) <- scat.species$species
scat.species$species <- NULL
scat.species[,1:22] <- sapply(scat.species[,1:22],as.integer)

sapply(scat.species, class)
#scat.species <- as.data.frame(scat.species)

scat.species$shannon <- diversity(scat.species, index="shannon") #calculate the indeces ONE BY ONE!
scat.species$simpson <- rarefy(scat.species, 2) - 1 #unbiased simpson

#scat.species$simpson <- diversity(scat.species, index="simpson")   #RAREFY! ,2) -1 (log)!!!!
 
scat.match <- match(speciesE$species, row.names(scat.species))
speciesE$shannon <- scat.species$shannon[scat.match]
speciesE$simpson <- scat.species$simpson[scat.match]

#s2 = read.delim('all_scats.txt')
#scat = s2[,33:54]
#scat[is.na(scat)] = 0
#scat$species <- paste(s2[,'genus'], s2[,'species'])

#Using SYNCSA for Rao's QE
require(SYNCSA)
scat.species$shannon <- NULL
scat.species$simpson <- NULL
rownames(scat.species) <- scat.species$species
scat.species$species <- NULL
r1 <- rao.diversity(scat.species)
scat.match <- match(speciesE$species, rownames(scat.species))
speciesE$raoS <- r1$Simpson[scat.match]


write.table(speciesE, "speciesE.txt", sep = "\t", row.names = FALSE)



#scat$sums <- rowSums(scat[1:22])


#We only have diet on 2 extinct species: Martes caurina AND Felis lybica


#-----------------------------
#manual calc

#n = pop
#N = sum(row)

for (i in 1:length(s2$species))	{
    assign(paste0("n",i), as.numeric(scat[i,1:22]))
}

N = rowSums(scat[1:22])


#p = n/N
for (i in 1:length(s2$species))	{
    assign(paste0("p",i),  get(paste0("n",i))/N[i])
}
  

for (i in 1:length(s2$species))	{
    scat$shannon[i,] <- diversity(get(paste0("n",i)))
}

H = -sum(p*log(p))  

rm(list=ls(pattern="n"))

#------------------------------------

```

##Camera trap abundance calc
```{r}

#Pop density approximated

#load up camera trap data
register$name <- paste(register[,'genus'], register[,'species'])
register_clean <- register[,c("name", "count", "count.per.day", "sample.no")]
register_clean$ratio <- round(register_clean$count/register_clean$count.per.day)
reg.species <- aggregate(. ~ name, data=register_clean, FUN=sum)
species.number <- as.data.frame(table(register_clean$name))
sp.match <- match(reg.species$name, species.number$Var1)
reg.species$number <- species.number$Freq[sp.match]
reg.species$rate <- (reg.species$count - reg.species$number) / reg.species$ratio


reg.match <- match(speciesE$species, reg.species$name)
speciesE$abundance <- reg.species$rate[reg.match]
speciesE$median.abundance <- NULL
write.table(speciesE, "speciesE.txt", sep = "\t", row.names = FALSE)



register_sample_co <- register_clean[,c(2:4)]
register_sample_co <- register_sample_co$count/register_sample_co$count.per.day

register_sample_co <- aggregate(. ~ sample.no, data=register_sample_co, FUN=sum)
sample.match <- match(register_sample_co$sample.no, register_clean$sample.no)
register_clean$total <- register_sample_co$count[sample.match]
register_clean$sample.no <- NULL

register_clean$density <- round(register_clean$count/register_clean$count.per.day)

```



##Wrangle, explore and investigate
```{r}
require(easystats)
require(psych)
describe_distribution(speciesE)
multi.hist(speciesE[,sapply(speciesE, is.numeric)], global = F) 



data_filter(speciesE, extant == 0 & is.na(diet.new)) #need to get the factors back with NA level to count it


require(lattice)
dotplot(factor(speciesE$extant) ~ log(speciesE$mass.g))
```


#Tree wrangle
```{r}
require(phytools)
require(BAMMtools)


treeF <- read.tree("Complete_phylogeny.nex")
rownames(speciesE) <- speciesE$species_

#Adding the 3 missing species (length = 0.5 on the branch - except for Conepatus = 0 from the mrca)

for (i in 1:length(tree))	{
tip <- "Martes_caurina"
sister <- "Martes_americana"

tree[i] <- c(bind.tip(tree[[i]],tip,where=which(tree[[i]]$tip.label==sister),
    position=0.5*tree[[i]]$edge.length[which(tree[[i]]$edge[,2]==
    which(tree[[i]]$tip.label==sister))])) 


tip <- "Conepatus_robustus"
sister <- "Conepatus_chinga"
sister2 <- "Conepatus_leuconotus" 

tree[i] <- c(bind.tip(tree[[i]],tip,where=getmrca(tree[[i]], sister, sister2),
    position=0*tree[[i]]$edge.length[getmrca(tree[[i]], sister, sister2)])) 

tip <- "Felis_lybica"
sister <- "Felis_silvestris"

tree[i] <- c(bind.tip(tree[[i]],tip,where=which(tree[[i]]$tip.label==sister),
    position=0.5*tree[[i]]$edge.length[which(tree[[i]]$edge[,2]==
    which(tree[[i]]$tip.label==sister))])) }


grep("Martes", tree[[7]]$tip.label, value = TRUE)


clean <- clean.data(speciesE, tree)
tree1 <- clean$tree[[1]] #we can use all of the trees in mcmcglmm
data1 <- clean$data

```


###*bad solution
```{r}
#adding missing species loop (bad solution with tree extraction!)

#extract all trees
treeNew <- list()
class(treeNew) <- "multiPhylo"


for(ntrees in 1:length(tree)) {
    assign(paste0("treeX", ntrees), tree[[ntrees]])
}

for (i in 1:length(tree))	{
tip <- "Martes_caurina"
sister <- "Martes_americana"

treeNew[i] <- c(bind.tip(get(paste0("treeX",i)),tip,where=which((get(paste0("treeX",i)))$tip.label==sister),
    position=0.5*(get(paste0("treeX",i)))$edge.length[which((get(paste0("treeX",i)))$edge[,2]==
    which((get(paste0("treeX",i)))$tip.label==sister))])) }
    
for(ntrees in 1:length(tree)) {
    assign(paste0("treeX", ntrees), treeNew[[ntrees]])
}


for (i in 1:length(tree))	{
tip <- "Felis_lybica"
sister <- "Felis_silvestris"
treeNew[i] <- c(bind.tip(get(paste0("treeX",i)),tip,where=which((get(paste0("treeX",i)))$tip.label==sister),
    position=0.5*(get(paste0("treeX",i)))$edge.length[which((get(paste0("treeX",i)))$edge[,2]==
    which((get(paste0("treeX",i)))$tip.label==sister))])) }

for(ntrees in 1:length(tree)) {
    assign(paste0("treeX", ntrees), treeNew[[ntrees]])
}

for (i in 1:length(tree))	{
tip <- "Conepatus_robustus"
sister <- "Conepatus_leuconotus"
treeNew[i] <- c(bind.tip(get(paste0("treeX",i)),tip,where=which((get(paste0("treeX",i)))$tip.label==sister),
    position=0.5*(get(paste0("treeX",i)))$edge.length[which((get(paste0("treeX",i)))$edge[,2]==
    which((get(paste0("treeX",i)))$tip.label==sister))])) }


rm(list=ls(pattern="treeX"))
```


#2. Load final tree and clean data
```{r}
require(phytools)
require(mulTree)
tree <- read.tree("CompletePhyCarn.txt")

clean <- clean.data(speciesE, tree)
tree1 <- clean$tree[[2]]
data <- clean$data

```

#Analyses
#~Functions
##~regress.multi.impute
```{r}
require(phylolm)

#Run phyloglm on multiple trees (with boot need to statrt from tree 2?!?)
regress.multi.impute <- function (multree, sets, model, from=1, to=length(tree), alpha=0.05, btol.v=10, boot.v=0) {
    a = 0 #counter for sig models
    b = 0
    z = matrix(ncol = length(colnames(dataX1)), nrow =sets*(to-from+1))
    total = sets*(to - from + 1)
    pb <- txtProgressBar(min = 0,      # Minimum value of the progress bar
                     max = total, # Maximum value of the progress bar
                     style = 3,    # Progress bar style (also available style = 1 and style = 2)
                     width = 50,   # Progress bar width. Defaults to getOption("width")
                     char = "=")   # Character used to create the bar
    #z <- NULL
    #z <- data.frame(matrix(ncol = 10, nrow = 10))
# gets the trees one by one, extracts the coeffs from the summary (as numeric), marks the sig p values 
# (see print.pval fun) and prints them - progress is displayed
for (i in from:to){
for (imps in 1:sets){
    temptree <- multree[[i]]
    s1 <- phyloglm(model, get(paste0("dataX", imps)), temptree, method = c("logistic_MPLE"), btol = btol.v, boot=boot.v)
    b = b + 1
    k = summary(s1)$coefficients[,3]
    #print(k)
    #print(length(k))
    z[b, 1 : length(k)] = k
    s2 <- as.data.frame(summary(s1)$coefficients[,"p.value"]) 
    asd <- as.numeric(s2[[1]])
    
    for(bar in 1:b) {
    setTxtProgressBar(pb, bar)}
    
    
    if(any(asd[-1]<alpha)){
        #print.pval(asd[-1]) 
        a = a+1
      cat("\n", "---------------", "\n", " N Sig models:", a, " out of", b, "", round((a/b)*100, 2), "% " ,"\n", "Models left:", (sets*(to - from + 1))-b, "\n") }
    
        #print(summary(s1)$coefficients[,4])
        #print (summary(s1))
        #print(z)
}
}
    close(pb)
    cat(paste0("\n Predictors: (", length(row.names(s2)[-1]),")"," + Intercept" ))
    cat(paste0("\n", " ", row.names(s2)[-1]), "\n")
    cat(paste0("\n ==================== \n", " Total models: ", to - from + 1, "x", sets, "=", (to - from + 1)*sets, " (trees x imputed sets)",  "\n Significant: ", a, "\n ", "p<", alpha," in ", round((a/((to - from + 1)*sets))*100, 2), "% of the tested sets", "\n ==================== \n"))
    zvalues <<- colMeans(abs(z), na.rm=T) # z or abs z?!
    #convert z values to p
    cat(paste0("\n P-values \n"))
    cat(paste0(row.names(s2)), "\n")
    pt(zvalues[1:length(row.names(s2))], ((length(row.names(dataX1)) - (length(row.names(s2))))), lower.tail = FALSE)
        }
#pt(zvalues, species - degs of F)
```

##RUN
```{r}

#function (multree, sets, model, from=1, to=length(tree), alpha=0.05, btol.v=10, boot.v=0)

regress.multi.impute(tree, 5, extant ~ mass.g, 1, 4)

```

##~print p and reg multi
```{r}

print.pval <- function(x, digits = 6) {
    sig_level <- cut(x,
                     breaks = c(-Inf, 0, 0.0001, 0.001, 0.01, 0.05, Inf),
                     labels = c("", "****", "***", "**", "*", ""))
    print(paste0(signif(x, digits = digits), sig_level), quote = FALSE)
    invisible(x)
}

#Run phyloglm on multiple trees (with boot need to statrt from tree 2?!?)
regress.multi <- function (multree, data, model, from=1, to=length(tree), alpha=0.05, btol.v=10, boot.v=0) {
    a = 0 #counter for sig models
# gets the trees one by one, extracts the coeffs from the summary (as numeric), marks the sig p values 
# (see print.pval fun) and prints them - progress is displayed
for (i in from:to){
    temptree <- multree[[i]]
    s1 <- phyloglm(model, data, temptree, method = c("logistic_MPLE"), btol = btol.v, boot=boot.v)
    s2 <- as.data.frame(summary(s1)$coefficients[,"p.value"]) 
    asd <- as.numeric(s2[[1]])
      if(any(asd[-1]<alpha)){
        print.pval(asd[-1]) 
        a = a+1
        cat(asd,"\n", " N Sig models:", a, " out of", i-from+1, "", round((a/(to-from))*100, 2), "% " ,"\n")
        }
        #print(summary(s1)$coefficients[,4])
        print (summary(s1))
                  }
    cat(paste0("\n Predictors: (", length(row.names(s2)[-1]),")" ))
    cat(paste0("\n", " ", row.names(s2)[-1]), "\n")
    cat(paste0("\n ==================== \n", " Total models: ", to - from + 1, "\n Significant: ", a, "\n ", "p<", alpha," ", round((a/(to-from+1))*100, 2), "% ", "\n ==================== \n"))
    }
regress.multi(tree, complete(imp, 2), extant ~ diet.cat, 1, 12)

```

## gls;phyloglm;rstanarm
```{r}

require(phylolm)
require(rstanarm)


#Non-phylo gls
a <- glm(speciesE$extant ~ speciesE$diet.1, family = binomial)
summary(a)

#RELEVEL for ancova relevel(diet.1, ref = "carnivore")

#timetree phy
tree2 <- read.newick("spcs.nwk")
clean.data(speciesE, tree)


aa <- phyloglm(extant ~ log(mass.g) + loctype, speciesE, tree) #scansorial+
aa <- phyloglm(extant ~ median.abundance + loctype, speciesE, tree) #terrestrial-
aa <- phyloglm(extant ~ diet.1, speciesE, tree) #carnivore+
aa <- phyloglm(extant ~ median.abundance + log(mass.g) + loctype, speciesE, tree) #median abd+, mass-

#phylacine phy
#treeF <- read.nexus("Complete_phylogeny.nex")
#rownames(speciesE) <- speciesE$species_
#clean <- clean.data(speciesE, treeF)
#tree1 <- clean$tree[1]$UNTITLED #we can use all of the trees in mcmcglmm
#data1 <- clean$data

#check the diet var!
#data1$diet.new <- as.factor(data1$diet.new)
#data1$diet.new <- factor(data1$diet.new, exclude="")

#USE THE MULTI.REGRESS

phyloglm(extant ~ log(BMR) + diet.1, data1, tree1, method = c("logistic_MPLE"), btol = 100, boot=10) # also only diet
phyloglm(extant ~ median.abundance + log(mass.g) + loctype, data1, tree1, method = c("logistic_MPLE"), btol = 30, boot = 100) # mass?

#rstanarm

fit_rstanarm <- stan_glm(
    extant ~ loctype,
    data = speciesE,
    family = "binomial"
)

summary (fit_rstanarm)
```


#3. PCA + imputation
```{r}

#Load data for pc, exclude chars
data.pc <- data[,3:15]
rownames(data.pc) <- data$species_name


#### Breaking down cat variables
options(na.action='na.pass') #return NA for missing data on cat vars

loctype <- model.matrix(~ 0 + a, data.frame(a = data.pc[,"loctype"])) # break down factors into 1/0
extant <- model.matrix(~ 0 + a, data.frame(a = data.pc[,"extant"]))
diet.cat <- model.matrix(~ 0 + a, data.frame(a = data.pc[,"diet.cat"]))
colnames(extant) <- c("Extinct0", "Extant1") #so they can be easily distinguished
data.pc <- cbind(data.pc, diet.cat, extant, loctype)

#remove old cats
data.pc$loctype <- NULL
data.pc$extant <- NULL



#Impute data.pc
require(mice)
imp <- mice(data.pc, 20, maxit = 5)

#run princomp on all imputed sets
#get loadings from all imputed sets in pcSET-component
#get scores from all imputed sets in scoresSET-component

for(imputedsets in 1: length(imp$imp[[1]])) {
  #extract imputed sets in separate dfs
  assign(paste0("dataX",imputedsets), complete(imp, imputedsets))
}

#extract diet.cat
for(imputedsets in 1: length(imp$imp[[1]])) {
  assign(paste0("diet.cat",imputedsets), get(paste0("dataX", imputedsets))[8]) #extract imputed diet cat
  assign(paste0("dataX",imputedsets), get(paste0("dataX",imputedsets))[-8])   #and remove it from the set before PCA
}


for(imputedsets in 1: length(imp$imp[[1]])) {
    assign(paste("PCA",imputedsets,sep=""), princomp(scale(get(paste("dataX",imputedsets, sep="")))))
for (component in 1 : length(PCA1$loadings[,1])) {     #number of components
      
  assign(paste("loadings",component, "-", imputedsets, sep=""), loadings(get(paste("PCA",imputedsets, sep="")))[,component]) 
  assign(paste("tempscores",component, "-", imputedsets, sep=""), (get(paste("PCA",imputedsets, sep=""))$scores)[,component]) 

  #put all loadings and scores PER set in separate objects
  #this works
  
  assign(paste0("aPC",component, sep=""), data.frame(do.call(cbind, mget(ls(pattern = paste0("\\bloadings", component, "\\b"))))))
  assign(paste0("Score",component, sep=""), data.frame(do.call(cbind, mget(ls(pattern = paste0("\\btempscores", component, "\\b"))))))
  
  }}






#get the means per component (both loadings and components)
for (component in 1 : length(PCA1$loadings[,1])) {
  
  assign(paste0("pc.means",component, sep=""), cbind(paste0("aPC",component, sep=""), rowMeans(get(paste0("aPC",component, sep="")))))
  assign(paste0("aPC",component, sep=""), cbind(get(paste0("pc.means",component, sep="")), get(paste0("aPC",component, sep=""))))
  
  assign(paste0("sc.means",component, sep=""), cbind(paste0("Score",component, sep=""), rowMeans(get(paste0("Score",component, sep="")))))
  assign(paste0("Score",component, sep=""), cbind(get(paste0("sc.means",component, sep="")), get(paste0("Score",component, sep=""))))
  
  assign(paste0("aPC",component, sep=""), get(paste0("aPC",component, sep=""))[-1])
  assign(paste0("Score",component, sep=""), get(paste0("Score",component, sep=""))[-1])
  
}

#change rowname from 2 to mean

for (i in paste0("aPC", 1:length(PCA1$loadings[,1]))){
  d=get(i)
  colnames(d)[1]='mean'
  assign(i,d)
}

for (i in paste0("Score", 1:length(PCA1$loadings[,1]))){
  d=get(i)
  colnames(d)[1]='mean'
  assign(i,d)
}
  
#get all the means in one df

MeanPC = aPC1[0:0]
for (i in paste0("aPC", 1:length(PCA1$loadings[,1]))) {
    m = get(i)[1]
    MeanPC[i] <- cbind(m)
    #assign(i, m)
}
MeanPC$mean <- NULL
MeanPC <- sapply(MeanPC, as.numeric)
MeanPC <- data.frame(MeanPC)
rownames(MeanPC) <- rownames(aPC1)

MeanS <- Score1[0:0]
for (i in paste0("Score", 1:length(PCA1$loadings[,1]))) {
    m = get(i)[1]
    MeanS[i] <- cbind(m)
    #assign(i, m)
}

MeanS <- sapply(MeanS, as.numeric)
MeanS <- data.frame(MeanS)
rownames(MeanS) <- data$species_name


#add to the dataX (original imputed sets)

#add loadings
for (i in paste0("dataX", 1:length(imp$imp[[1]]))) {
    assign(i, cbind(get(i), MeanLoadings))
}


for (i in paste0("dataX", 1:length(imp$imp[[1]]))) {

    assign(i, get(i)[-c(11:23)])                    #remove broken down cats
    assign(i, cbind(get(i), data[,c(5,8)]))         #add old cats
}


#add back the imputed diet.cat
for(imputedsets in 1: length(imp$imp[[1]])) {
    assign(paste0("dataX", imputedsets), cbind(get(paste0("dataX", imputedsets)), get(paste0("diet.cat",imputedsets))))  #add diet.cat back
}
  
  
#clean up    

rm(list=ls(pattern="pc.means"))
rm(list=ls(pattern="sc.means"))
rm(list=ls(patter= "PCA"))
#rm(list=ls(pattern="dataX"))
rm(list=ls(pattern="aPC"))
rm(list=ls(pattern="diet.cat"))
rm(list=ls(pattern="loadings"))
rm(list=ls(patter= "tempscores"))
rm(list=ls(patter= "Score"))


```

#~Plots
```{r}

#clean <- clean.data(data, tree1)
#tree1 <- clean$tree
#data <- clean$data1


#choose cols to plot
colnames(data)

#Get them by number
#nameCol <- data.frame(colnames(data))
#cols <- nameCol[c(4,6,8),]

#Get them by name
cols <- c("loctype","diet.new", "extant")

#Fetch only the desired columns in a new df and give it row names
data1 <- data[c(cols)]

#convert vars to what they need to be - factor, numeric etc (if need be) (factors with NA level)
data1[cols] <- lapply(data1[cols], factor, exclude="")

#check if it worked
levels(data1[,1])
sapply(data1, class)

#get 1 tree
tree1 <- tree[[2]]

require(ggtree)
require(ggplot2)
require(wesanderson)

pal <- wes_palette("Darjeeling1", 1110, type = "continuous")


#tree1 <- drop.tip(tree1, "Felis_catus")


#Circular tree with traits from data1
circ <- ggtree(tree1, layout = "circular")
p1 <- gheatmap(circ, data1, offset=.8, width=.2,
               colnames_angle=95, colnames_offset_y = .25) + scale_fill_viridis_d(option="D", name="discrete\nvalue")

#Visualise traits per cat
p2 <- ggplot(data1, aes(x=extant, y=tree1$tip.label, width=.5)) + 
    geom_tile(aes(fill=`extant`)) + scale_fill_viridis_d() + 
    theme_minimal() + xlab(NULL) + ylab(NULL)

#+ geom_tippoint(aes(color = c("a", "b")))

#circ tree with offset
circ <- ggtree(tree1, layout = "circular") + geom_tiplab(offset=15)
p1 <- gheatmap(circ, data1, offset=.8, width=.2,
               colnames_angle=95, colnames_offset_y = .25) + scale_fill_viridis_d(option="D", name="discrete\nvalue")

#Tree with offset
g <- ggtree(tree1) + geom_tiplab(align=TRUE, offset=15) + hexpand(.5)

#Plot the heatmap
p5 <- gheatmap(g, data[cols], offset=.8, width=.2,colnames_angle=95, colnames_offset_y = .25, colnames_position = "top")


#grep("robustus", b, value = TRUE)


#heatmap using ggtree

p11 <- ggtree(tree[[1]]) + 
    xlim(0, 150) +
    geom_tiplab(size=4, offset = 35)

p21 <-  gheatmap (p11, data[,c("diet.1", "diet.cat", "diet.new")], offset=-0.5, width=0.5, low=(values=pal[1]), high=(values=pal[2]), colnames_position = "top", font.size=4) + scale_fill_manual(values = pal)

#phylo.heatmap(tree,X,standardize=TRUE)


library(aplot)
p21 %>% insert_left(p5) %>% insert_right(p1, width=.5) 




library(ggridges)
ggplot(speciesE, aes(x = brain, y = extant, fill = stat(x))) +
    geom_density_ridges_gradient() +
    scale_fill_viridis_c(name = "Depth", option = "C")


ggplot(speciesE, aes(x = litter, y = loctype, fill = stat(x))) +
    geom_density_ridges_gradient() +
    scale_fill_viridis_c(name = "Depth", option = "C") +
    coord_cartesian(clip = "off") + # To avoid cut off
    theme_minimal()

ggplot(speciesE, aes(x = litter, y = loctype, fill = stat(x))) +
    geom_density_ridges_gradient(stat="binline") +
    scale_fill_viridis_c(name = "Depth", option = "C") +
    coord_cartesian(clip = "off") + # To avoid cut off
    theme_minimal()


speciesE %>%
     mutate(diet.cat = factor(diet.cat)) %>%
     ggplot() +
     geom_boxplot(aes(x = diet.cat, y = log(mass.g)))

ggplot(speciesE) + 
    geom_density_ridges(aes(x = log(mass.g), y = diet.cat, group = interaction(diet.cat, extant),fill = extant), alpha = 0.7)

```

##~PCA plots
```{r}


PCAcolors <- c("#66c2a5","#fc8d62","#8da0cb")[as.integer(data$extant)]

PCAscores <- MeanS
PCAloadings <- MeanPC    #[1:10,]
components <- 4

par(mfrow=c(4,2)) 
for (i in 1:components){
a=i #can fix to comp N
b=i+1
plot(PCAscores[,c(a,b)],  # x and y data
     pch=21,           # point shape
     col=PCAcolors,    # point border color
     bg=PCAcolors,     # point color
     cex=1.5,          # point size
     main="Scores"     # title of plot
)
legend("topright",                                # position of legend
       legend=levels(data$extant),                       # legend display
       pch=21,                                    # point shape
       pt.bg=c("#66c2a5","#fc8d62","#8da0cb"),    # point colors
       pt.cex=1.5,                                # point size
       col = c("#66c2a5","#fc8d62","#8da0cb")    # point border color
)
plot(PCAloadings[,c(a,b)],   # x and y data
     pch=21,              # point shape
     bg="black",          # point color
     cex=1,               # point size
     main="Loadings"      # title of plot
)
text(PCAloadings[,c(a,b)],             # sets position of labels
     labels=rownames(PCAloadings)   # print labels
)}

dev.new()
plot(PCAscores[,1:4],  # x and y data
     pch=21,           # point shape
     col=PCAcolors,    # point border color
     bg=PCAcolors,     # point color
     cex=1.5,          # point size
     main="Scores"     # title of plot
)

```





###############3


species -> register
register <- sample
registerX$species_full <- trimws(registerX$species_full, which = c("right"))


rnum <- array(dim=245, dimnames = list(as.character(nos)), data = 1:245) 



for (i in nos){
    w = which(registerX[,'sample.no'] == i)
    samplesX[rnum[as.character(i)], "countT"] = sum(registerX[w, 'count'])
}


Nowhere to run nowhere to hide! :D


#Misingness (and imputation)
```{r}
require(naniar)
require(mice)

gg_miss_var(speciesE)
vis_miss(speciesE)

imp <- mice(data[,3:15], 20, maxit = 10)

plot(imp)
densityplot(imp)
stripplot(imp, pch = 20, cex = 1.2)

complete(imp, 2)
```

#Path
```{r}
library(phylopath)

models <- define_model_set(
    one   = c(mass.g ~ extant, brain ~ extant),
    two = c(extant ~ mass.g, brain ~ mass.g),
    #three   = c(extant ~ brain, brain ~ mass.g),
    #four = c(extant ~ mass.g + brain),
    five = c(extant ~ mass.g, extant ~ brain))

plot_model_set(models)


result <- phylo_path(models, data = data1, tree = tree[[1]], model = 'lambda', method = "logistic_MPLE")
s <- summary(result)
plot(s)
best_model <- best(result)
plot(best_model)
average_model <- average(result)
plot(average_model, algorithm = 'mds', curvature = 0.1)
```